\documentclass{homework}
\author{Wiyninger, Caleb}  % Uncomment with your name
\class{MATH 3503: Tashfeen's Discrete Mathematics}
\date{\today}  % Uncomment with your semester
\title{Homework 1}
\address{
  Computer Science, %
  Petree College of Arts \& Sciences, %
  Oklahoma City University
}

\begin{document} \maketitle

\question Please read chapters 0 and 1 of Chartrand et al. 
and write a couple sentences about a /example/concept that you found 
difficult or interesting and why?

\begin{sol}

  I enjoyed the section where it described how the google page rank 
  algorithm works. I've learned about how it works before but having it 
  described in terms of discrete mathematics connected it to interesting 
  things ive learned already. \bigbreak

  I struggled a little with the implication word problems. 
  I think I understand the concept fine, but determining what part of 
  the sentence corresponds to what was a little confusing

\end{sol}                              

\question\label{a} Let $P, Q$ and $R$ be statements. Determine whether the following is true.
\[
    P \oplus (Q \oplus R) \is (P \oplus Q) \oplus R. \quad \text{Use a truth table.}
\]

  \begin{sol}
\tbl<tb1>{Truth table for question \ref{a}.} {
  $P$ & $Q$ & $R$ & $P \oplus (Q \oplus R)$ & $(P \oplus Q) \oplus R$ \\
  F & F & T & T & T \\
  F & T & T & F & F \\
  F & F & F & F & F \\
  F & T & F & T & T \\
  T & F & T & F & F \\
  T & T & T & T & T \\
  T & F & F & T & T \\
  T & T & F & F & F \\             
}

$P \oplus (Q \oplus R) \is (P \oplus Q) \oplus R$ is true

\end{sol}

\question\label{b} Let $P, Q$ and $R$ be statements. Determine whether the following is true.
\[
    P \vee (Q \oplus R) \is (P \vee Q) \oplus (P \vee R). \quad \text{Use a truth table.}
\]

% type your solution here
 \begin{sol}
\tbl<tb2>{Truth table for question \ref{b}.} {
  $P$ & $Q$ & $R$ & $P \vee (Q \oplus R)$ & $(P \vee Q) \oplus (P \vee R)$ \\
  T & T & T & T & F \\
  T & T & F & T & F \\
  T & F & T & T & F \\
  T & F & F & T & F \\
  F & T & T & F & F \\
  F & T & F & T & T \\
  F & F & T & T & T \\
  F & F & F & F & F \\
 }
 
 $P \vee (Q \oplus R) \is (P \vee Q) \oplus (P \vee R)$ is not true
\end{sol}

\question For an integer $n$, consider the open sentences
\[
    P(n): n(n + 1)(2n + 1)/6\text{ is even.} \quad Q(n): (n + 1)^2 (n + 2)^2 /4\text{ is even.}
\]
Determine whether the biconditionals $P(1) \iff Q(1)$ and $P(2) \iff Q(2)$ are true or false.

% type your solution here
\begin{sol}
  $P(1) = 1$ is even = false \smallbreak
  $Q(1) = 9$ is even = false\smallbreak
  $P(1) \iff Q(1) = true$ \bigbreak
  $P(2) = 5$ is even = false\smallbreak
  $Q(2) = 36$ is even = true \smallbreak
  $P(2) \iff Q(2) = false$ \smallbreak
\end{sol}

\question\label{c} For statements $P$ and $Q$, determine whether the compound statement
\[
    (P \wedge (\sim Q)) \Ra (P \vee Q)
\]
is a tautology, a contradiction or neither.

% type your solution here
\begin{sol}
  \tbl<tb3>{Truth table for question \ref{c}.} {
  $P$ & $Q$  & $(P \wedge (\sim Q))$ & $(P \vee Q)$ & $(P \wedge (\sim Q)) \Ra (P \vee Q)$  \\
  T & T & F & T & T \\
  F & F & F & F & T \\
  T & F & T & T & T \\
  F & T & F & T & T \\
 }
  $(P \wedge (\sim Q)) \Ra (P \vee Q)$ is a tautology
\end{sol}

\question Each of the following statements is an implication $P \Ra Q$. For each statement, indicate what $P$ and $Q$ are.
\begin{enumerate}[label=(\alph*)]
    \item I'm going to my class reunion only if I lose weight.

    \item To win a free \$20 gift certificate, I must spend \$100 at the store.

    \item To win the game, it is necessary that we score a touchdown.

    \item It is necessary to do research to be promoted to professor.

    \item I'll get an A on this exam if I'm lucky.
    
    \item All I need is a B on the final exam to get an A in the course. 
  \end{enumerate}
  \begin{sol}
    - a. \smallbreak
    P = Going to class reunion \smallbreak
    Q = If i lose weight \bigbreak
    - b. \smallbreak
    P = Won certificate \smallbreak
    Q = Spend 100 dollars \bigbreak
    - c. \smallbreak
    P = Won the game \smallbreak
    Q = Scored touchdown \bigbreak
    - d. \smallbreak
    P = Promoted to professor \smallbreak
    Q = Did research \bigbreak
    - e. \smallbreak
    P = Got A \smallbreak
    Q = Is lucky \bigbreak
    - f. \smallbreak
    P = Got a B on the exam \smallbreak
    Q = Got an A in the course \smallbreak
  \end{sol}


\end{document}
